---
title: "Cake Ingredient Analysis"
output: html_notebook
---

```{r}
library(tidyverse)
library(readr)
library(janitor)
library(skimr)
library(stringr)
library(here)
```


```{r}
# Read in the data
cake_recipes <- read_csv(here("data/raw_data/cake-ingredients-1961.csv"))
cake_codes <- read_csv(here("data/raw_data/cake_ingredient_code.csv"))
```

# Cake Ingredient Analysis

## The Raw Data

The *raw data* came in the form of two .csv files. Table 1 (cake_ingredients_1961.csv)  contains all the different types of cake with all of the `ingredient codes` as column headers. This file was in __wide format__ and filled with **NA values**. Table 2 (cake_ingredient_code) contains a list of ingredients and their correspongding _ingredient code_, as well as a measure column. The only thing missing from this table was the `measure` value for "Sour cream cup".

## Assumptions

My first assumption was that I would have to change the format of Table 1 to __long format__. This would make it easier to see which ingredients were required for each cake. I could see there were many _NA values_ in this table and I was assuming that these ingredients were not required for the corresponding cake, as oppose to *missing information*. The brief states that the full ingredient name is required and not just the `ingredient code`. So I thought the best way to approach this is to **join** the tables using the `ingredient code`. 

Another assumption I made about table 2 was that the *missing information* in the `measure` column was meant to be "cup". I assumed that the `ingredient` listed was meant to be "Sour cream" and the `measure` was meant to be "cup". Rather than the ingredient being "Sour cream cup".


## Cleaning Proceess Overview

1. My first step was to pivot the table in to __long format__. I chose to do this rather than clean the column headers as I knew that it would be easier keeping them in upper-case to join the tables.

2. I then decided that was the best time to get rid of _NA values_ so i used _drop_na_
```{r}
cake_recipes_long <- cake_recipes %>%
     pivot_longer(cols = ("AE":"ZH"),
                names_to = "code",
                values_to = "quantity")

cake_recipes_dropna <- cake_recipes_long %>%
  drop_na(quantity)
```

3. Now seemed like the time to `join` the tables, clean the header names and select only the columns I required (get rid of the _ingredient code_).
```{r}
 cake_recipes_full <- cake_recipes_dropna %>%
  left_join(cake_codes, by = "code")

cake_recipes_full_clean <- cake_recipes_full %>%
  clean_names()

cake_recipe_clean <- cake_recipes_full_clean %>%
  select(cake, ingredient, measure, quantity)
cake_recipe_clean
```

4. The last step in my cleaning process was dealing with the *missing information*. This was in the `measure` column corresponding to "Sour cream cup". As previously mentioned I removed the "cup" from the `ingredient` column and added it to the `measure` column.
```{r}
cake_recipe_clean_sour <- cake_recipe_clean %>%
  mutate(measure = coalesce(measure, "cup"))

cake_recipe_clean_cup_remove <- cake_recipe_clean_sour %>%
  mutate(ingredient = str_remove(ingredient, " cup"))

cakes_cleaned <- cake_recipe_clean_cup_remove
```

## Task Analysis

Read in cleaned data
```{r}
cakes <- read_csv(here("data/clean_data/cakes_clean.csv"))
```

1. Which cake has the most cocoa in it?
```{r}
most_cocoa_cake <- cakes %>%
  filter(ingredient == "Cocoa") %>%
  slice_max(quantity)
```
- The `One Bowl Chocolate` cake has the most `Cocoa` in it by far.


2. For sponge cake, how many cups of ingredients are used in total?
```{r}
sponge_cake_test <- cakes %>% 
  filter(cake == "Sponge",
         measure == "cup") %>% 
  summarise(total_quantity = sum(quantity))
```
- Filter by _Sponge_ and _cup_ and then find the total of those values, which gives us 3.5 cups.


3. How many ingredients are measured in teaspoons?
```{r}
n_in_teaspoons <- cakes %>% 
  filter(measure == "teaspoon") %>% 
  distinct(ingredient) %>% 
  summarise(n_ingredients_teaspoon = n())
```
- Filter by _teaspoon_ and use the **distinct()** function to find all of the individual ingredients that are measured in teaspoons.


4. Which cake has the most unique ingredients?
```{r}
most_unique_ingred <- cakes %>% 
  group_by(cake) %>% 
  summarise(total_no_ingred = n()) %>% 
  arrange(desc(total_no_ingred)) %>% 
  head(2)
```
- On finding the sum of the ingredients for each cake I discovered that both the `Babas au Rhum` and `One Bowl Chocolate` both contained 11 ingredients. Which was the highest number of ingredients of all of the cakes.


5. Which ingredients are used only once?
```{r}
ingedients_once <- cakes %>% 
  group_by(ingredient) %>% 
  mutate(times_used = n()) %>% 
  filter(times_used == 1) %>% 
  select(ingredient, times_used)
```
- **Grouped by** ingredients and then counted how many times each one occurs. I then filtered to show all of the ingredients which occured only once.