---
title: "Candy Overview"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(here)
library(readxl)
library(lubridate)
library(stringr)
```

```{r}
# Read in the cleaned data
bb_candy_bind_clean <- read_csv(here("data/clean_data/bb_candy_bind_clean.csv"))
```

# Boing Boing Candy Analysis

## The Raw Data

The *raw data* came in the form of 3 .xlsx files. These contained data from 3 years worth of trick or treating information. They were all in __wide format__ and filled with **NA values**. There was a lot of data that was not required for the analysis. There were a couple of columns with problematic data which required a lot of cleaning. In particular the *country* column in the 2016 and 2017 table had many different spellings of the country names required. Also the age tables were of type __character__ and had a lot of values which sat outside my chosen range.

## Assumptions

My first assumption was that I would have to trim off the multiple columns which were not required for the analysis. I also knew that I would have to __pivot__ the table/tables to **long format** at some point. This massively reduced the number of columns and made the data easier to read. Also, I knew I had to __bind__ the tables. I decided that I would do a lot of the **cleaning** of the column headers before binding and then clean the information in the columns after, such as country names and gender.

## Cleaning Proceess Overview

I have four .R scripts containing all of my cleaning data. I have one for each one of the years (2015, 2016, 2017), then a final cleaning script with all of the data in one table (bb_candy_bind_clean.csv). At the end of this script I wrote the final cleaned table to a `.csv` file ready for analysis. Here is an overview of my cleaning steps.

- Read in the .xlsx file and assign it.
- Use the `clean_names()` function to tidy up the column headers.
- Use `rename()` and `str_remove()` to further tidy up column names. 
- Change the column __type__ using `mutate()` and `as.numeric()`.
- Amend the date as we only required the **year** for analysis.
- Remove all of the columns which we do not require.

## Task Analysis

1. What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)
```{r}
n_candy_ratings <- bb_candy_bind_clean %>% 
  summarise(n())
n_candy_ratings
```
- Every row in the dataset represents a rating as the __NA values__ have already been removed. Therefore the total number of ratings is just the number of rows, which is 648842.


2. What was the average age of people who are going out trick or treating?
```{r}
avg_age_t_or_t <- bb_candy_bind_clean %>% 
  filter(going_trick_or_treating == "Yes") %>% 
  drop_na(age) %>% 
  summarise(mean_age = round(mean(age), 2))
avg_age_t_or_t
```
- **Filtered** to only show the people who went trick or treating. Dropped the __NA values__ as they will hinder the `summarise`. Found the `mean` of all of the ages left and **rounded** to 2 decimal places, giving us 35.28.


3. What was the average age of people who are not going trick or treating?
```{r}
avg_age_not_t_or_t <- bb_candy_bind_clean %>% 
  filter(going_trick_or_treating == "No") %>% 
  drop_na(age) %>% 
  summarise(round(mean(age), 2))
avg_age_not_t_or_t
```
- Similar process to the previous question. The `mean` age of people not going trick or treating is 39.24 which is what we would expect as usually younger people go trick or treating.


4. For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r}
n_of_each_rating <- bb_candy_bind_clean %>% 
  group_by(rating) %>% 
  summarise(n_ratings = n())
n_of_each_rating
```
- **Grouped by** `rating` and summarised a count of the number of each rating. Surprisingly more candies brought __despair__ than **joy**.


5. How many people rated Starburst as despair?
```{r}
starburst_despair <- bb_candy_bind_clean %>% 
  filter(candy == "starburst",
         rating == "despair") %>% 
  summarise(n_of_starburst_despair = n())
starburst_despair
```
- **Filtered** by `candy` and `rating` then **summarised** with a count of the people who rated starburst as despair. This number was 1990.


- For the next three questions, count despair as -1, joy as +1, and meh as 0.
```{r}
bb_candy_bind_rating_system <- bb_candy_bind_clean %>% 
  mutate(rating_no = case_when(
    rating == "joy" ~ 1,
    rating == "despair" ~ -1,
    rating == "meh" ~ 0
  ))
bb_candy_bind_rating_system
```
- First I created a new column called `rating_no` which gave a value to each of the 3 different ratings. 


6. What was the most popular candy bar by this rating system for each gender in the dataset ?
```{r}
bb_candy_all_gender_favs <- bb_candy_bind_rating_system %>% 
  group_by(candy, gender) %>%
  summarise(sum_rating = sum(rating_no)) %>% 
  group_by(gender) %>% 
  slice_max(sum_rating)
bb_candy_all_gender_favs
```
- I initially **grouped_by** `candy` and `gender` to find the `sum_rating` of the `rating_no`. I then had to group by year again and use **slice_max** to find the result above. This shows that _reese_s_peanut_butter_cups_ seems to be a favourite.


7. What was the most popular candy bar in each year?
```{r}
bb_candy_most_pop_by_year <- bb_candy_bind_rating_system %>% 
  group_by(candy, year) %>%
  summarise(sum_rating = sum(rating_no)) %>% 
  group_by(year) %>% 
  slice_max(sum_rating)
bb_candy_most_pop_by_year
```
- Similar to the previous analysis, I **grouped_by** `candy` and `year` to find the `sum_rating` of the `rating_no`. I then grouped again by year and used **slice_max()** to give me the most popular candy bar from each year.


8. What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?
```{r}
bb_candy_country_favs <- bb_candy_bind_rating_system %>% 
  group_by(candy, country) %>%
  summarise(sum_rating = sum(rating_no)) %>% 
  group_by(country) %>% 
  slice_max(sum_rating)
bb_candy_country_favs
```
- Again, similar to the previous analysis. I had to **group_by** twice to give me the result above. Interestingly enough this shows a three way tie between `lindt_truffle`, `rolos` and `toblerone` as the uk's favourite candy!










